Paper

### 2018-01-22 Paper.Rmd
...

pander( 
  as.data.frame(all.info %>% group_by(Status) %>% summarise(
  'Total Patients'=length(unique(Patient)),
  'Total Samples'=sum(total.samples),
  'Median Years'=paste(median(end.year-start.year),'+/-',round(sd(end.year-start.year),1)),
  'Median Endoscopies'=paste(median(total.endos),'+/-',round(sd(total.endos),1)),
  'Median age'=paste(median(age.diagnosis),'+/-',round(sd(age.diagnosis),1)),
  'gender F:M'=paste(table(gender),collapse=':')
))
, caption='Full cohort')

@@

@@

```{r echo=F, warning=F, message=F, eval=T, fig.width=12, fig.height=8}

@@

globalCN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms'), labeller(Status=n.status, chr=label_value))
cnLegend = get.legend(globalCN.plot)
#globalCN.plot = globalCN.plot + theme(legend.position = 'none')

## No HGD
`%nin%` = Negate(`%in%`)

nhgd.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology %nin% c('HGD','IMC'))$variable)), ')', sep=''),
              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology %nin% c('HGD','IMC'))$variable)),')', sep=''))
globalnHGDCN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms' & Pathology %nin% c('HGD','IMC')), labeller(Status=nhgd.status, chr=label_value)) + labs(title='Excluding HGD/IMC')

@@

globalCN.plot 
globalnHGDCN.plot
globalBECN.plot

dfpts = subset(globalCNdf, Patient %in% c('57','76') & seg.type != 'arms')
#dfpts = dfpts %>% group_by(Patient, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )

pt.status <- c('Non-Progressor'=paste('Patient 57 (n=',length(unique(subset(dfpts, Patient == 57)$variable)), ')', sep=''),
              'Progressor'=paste('Patient 76 (n=',length(unique(subset(dfpts, Patient == 76)$variable)),')', sep=''))
genomePtCN.plot = cn.mtn.plot(dfpts, labeller(Status = pt.status))

dfpts = subset(globalCNdf, Patient %in% c('57','76') & Pathology == 'BE' & seg.type != 'arms')
#dfpts = dfpts %>% group_by(Patient, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )

ptbe.status <- c('Non-Progressor'=paste('Patient 57 (n=',length(unique(subset(dfpts, Patient == 57)$variable)), ')', sep=''),
              'Progressor'=paste('Patient 76 (n=',length(unique(subset(dfpts, Patient == 76)$variable)),')', sep=''))
genomePtBECN.plot = cn.mtn.plot(dfpts,  labeller(Status = ptbe.status )) + labs(title='NDBE samples only')

genomePtCN.plot 
genomePtBECN.plot

@@

a = sample( subset(subset(patient.info, Patient == 57),  Endoscopy.Year == min(Endoscopy.Year) )$Samplename, 1)
b = sample( subset(subset(patient.info, Patient == 76), Endoscopy.Year == min(Endoscopy.Year) & Pathology %nin% c('HGD','IMC'))$Samplename, 1)

ind.status <- c('Non-Progressor'='Patient 57 (n=1)',
              'Progressor'='Patient 76 (n=1), excl. HGD/IMC')
cn.mtn.plot(subset(globalCNdf, variable %in% c(a,b) & seg.type != 'arms'), labeller(Status = ind.status), type='bar') + labs(title="Individual samples")
```

@@

ch$ymax = ifelse(ch$coef > 0, max(genomePtCN.plot$data$value), min(genomePtCN.plot$data$value))
chh = chh %>% rowwise() %>% mutate('max'=custom.max(coef, hazards, ymax))
genomePtCN.plot +  geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F)

genomePtBECN.plot +  geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F)

@@

```{r prediction_cutoff, echo=F, message=F, warning=F, fig.height=5, fig.width=5}
preds = unlist(lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], function(df) df$Prediction))
np.preds = unlist(lapply(pg.samp[subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID], function(df) df$Prediction))

roc50 = pROC::roc(as.integer(c(preds, np.preds) > 0.5), c(rep(1, length(preds)), rep(0, length(np.preds))), main='Progressors & Non-progressors, cutoff 0.5')
roc50$model = 'CN Model 0.5'

preds = do.call(rbind.data.frame, lapply(pg.samp, function(df) df[c('Status','Prediction')]))
roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds')
roc$model = 'CN Model'

@@

ggplot() + geom_line(data=df, aes(specificity, sensitivity)) + scale_x_reverse() + 
  geom_point(data=thresholds, aes(x=specificity,y=sensitivity,color=label), size=3,shape=8, show.legend = T) +
  geom_label_repel(data=thresholds, aes(x=specificity,y=sensitivity,fill=label,label=
    paste('Threshold ', round(threshold,2), '\n(FPR ', round((1-specificity),2)*100, '%, TPR ', round(sensitivity,2)*100,'%)\nAUC:',round(roc$auc, 2)*100, '%', sep='')),show.legend = F, color='white',nudge_x=0.05) + scale_fill_manual(values=brewer.pal(3,'Dark2')) + scale_color_manual(values=brewer.pal(3,'Dark2')) +
  labs(title='CN Model ROC', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme

roc.plot(roc, 'Probability threshold calculated by ROC')

@@

```{r echo=F, fig.height=6, fig.width=8}
roc$model = "Per Sample"
multi.roc.plot(list(roc, rocEndoM, rocEndoA, rocPtM, rocPtA,rocFirstEndo,rocLastEndo), 'ROC Comparisons')
```

@@

get.auc.at(p53roc)
get.auc.at(hgdpathroc)
get.auc.at(pathroc)
get.auc.at(roc)

@@

rocList = list(roc,roc50, pathroc,hgdpathroc,p53roc)

dodge = position_dodge(width=0.9)
ggplot(melt(subset(m, Measure == 'AUC'), measure.vars='value'), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=Measure)) + 
  geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='black' ) + 
  geom_text(aes(label=paste(round(value,2)*100,'%',sep=''), y=0, vjust=-1 ), position=dodge, color='white')  +
  scale_fill_manual(values=c('purple3','dodgerblue3','firebrick3'), name='') + plot.theme + 
  theme(legend.position = 'none') + labs(x='', y='Percent', title='ROC Comparison') 
```

***

__All predictions from here onwards are performed using the `r cutoff` cutoff__

***

@@

```{r, echo=F, warning=F, message=F, fig.height=12, fig.width=16}

mtn.spatial<-function(df, b, limits, title='', pal=NULL) {
  if (is.null(pal))
    pal = RColorBrewer::brewer.pal(length(limits), 'Paired')

  df = merge(df, b[,c('Samplename','Endoscopy.Year','months.before.final','Block','nl')], by.x='variable',by.y='Samplename')
  df$nl = factor(df$nl, ordered = T)

  #pal = pal[which(limits %in% levels(df$nl))]

  ggplot(df, aes(x=chr.length)) + facet_grid(Endoscopy.Year~chr, scales='free_x', space='free_x') +
    geom_rect(aes(xmin=start, xmax=end, ymin=0, ymax=mean.value, fill=nl), show.legend=T) + 
    scale_fill_manual(values=pal, limits=limits,  name='OGJ Dist.') +
    theme(axis.text.x = element_blank()) +
    plot.theme + labs(title=title)
}

ogjLimits = levels(factor(back$nl, ordered = T))

p1 = mtn.spatial(subset(globalCNdf, Patient == 48 & seg.type != 'arms'), back, ogjLimits, 'Patient 48, (P)')
ogjLegend = get.legend(p1)
p1+theme(legend.position = 'none')

mtn.spatial(subset(globalCNdf, Patient == 60 & seg.type != 'arms'), back,ogjLimits,'Patient 60 (P)' )+theme(legend.position = 'none')

mtn.spatial(subset(globalCNdf, Patient == 3 & seg.type != 'arms'), back,ogjLimits,'Patient 3 (NP)' )+theme(legend.position = 'none')

mtn.spatial(subset(globalCNdf, Patient == 15 & seg.type != 'arms'), back,ogjLimits,'Patient 15 (NP)' )+theme(legend.position = 'none')

@@

```{r, echo=F, warning=F, message=F, fig.height=8, fig.width=9}
riskPath = rbind( cbind.data.frame( table(subset(back, Status == 'P', select=c('Pathology','Risk'))), 'Status'='Progressor' ), 
                  cbind.data.frame( table(subset(back, Status == 'NP', select=c('Pathology','Risk'))), 'Status'='Non-Progressor' ))

@@

riskPath = merge(riskPath, pathTotal)
riskPath$ratio = with(riskPath, round(Freq/nPath, 2))
riskPath$Risk = factor(riskPath$Risk, levels=c('Low','Moderate','High'), ordered = T)

p53Path = cbind.data.frame(table(subset(back, Status == 'P', select=c('Pathology','p53.Status'))), 'Status'='p53 IHC')
p53Path = merge( subset(back, Status == 'P' & !is.na(p53.Status)) %>% group_by(Pathology) %>% summarise( 'nPath'=length(Pathology)), p53Path)
p53Path$Risk = with(p53Path, ifelse(p53.Status == 1, 'High','Low'))
p53Path$ratio = with(p53Path, round(Freq/nPath, 2))

@@

ggplot(riskPath, aes(Pathology, ratio*100)) + geom_bar(aes(group=Risk, fill=Risk),stat='identity', position=position_stack()) + 
  scale_fill_manual(values=riskCols) + 
  geom_text(data=pathTotal, aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101), position=position_stack(), size=4) +
  geom_text(data=subset(riskPath, ratio>0), aes(group=Risk,label=paste(ratio*100,'%',sep='')), position=position_stack(vjust = 0.5), size=5) +
  facet_grid(~Status) + plot.theme + labs(title='Samples predicted by pathology', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')

@@

```{r p53IHC, echo=F, warning=F, message=F, fig.height=6, fig.width=8}

p53Path$Risk[p53Path$Risk == 'High'] = 'Aberrant'
p53Path$Risk[p53Path$Risk == 'Low'] = 'Normal'

col = brewer.pal(11, "PRGn")[c(3,9)]
ggplot(p53Path, aes(Pathology, ratio*100, group=Pathology, fill=Risk)) + geom_bar(stat='identity', position=position_stack()) + 
  scale_fill_manual(values=col) + 
  geom_text(aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101)) +
  geom_text(aes(label=paste(ratio*100,'%',sep='')), position=position_stack(vjust = 0.5), size=5) +
  plot.theme + labs(title='Samples predicted by aberrant p53 IHC', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')

@@

```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=15 }
back$nl = factor(back$nl, ordered=T)
back$Risk = factor(back$Risk, levels=c('Low','Moderate','High'), ordered = T)
back$Patient = factor(as.numeric(back$Patient), ordered = T)

pred.tiles(subset(back, Patient %in% c(6)), ncol=1, cols=myPal, probs=T) + labs(title='') + theme(legend.position = 'none')
pred.tiles(subset(back, Patient %in% c(48)), ncol=1, cols=myPal, probs=T) + labs(title='') + theme(legend.position = 'none')
pred.tiles(subset(back, Patient %in% c(33)), ncol=1, cols=myPal, probs=T) + labs(title='')+ theme(legend.position = 'none')
pred.tiles(subset(back, Patient %in% c(89)), ncol=1, cols=myPal, probs=T) + labs(title='')+ theme(legend.position = 'none')

grid.arrange(
  pred.tiles(subset(back, Patient %in% c(13,20,34,56,60,48) & Status == 'P'), ncol=1, cols=myPal, probs=T) + labs(title='Progressors'),
  pred.tiles(subset(back, Patient %in% c(15,16,21,3,45,19) & Status == 'NP'), ncol=1, cols=myPal, probs=T) + labs(title='Non-Progressors'),
ncol=2)

@@

head(back.v)

@@

roc$model = 'Discovery'
multi.roc.plot(list(roc, vroc), title='ROC')

>>>>

pander(
  as.data.frame(all.info %>% summarise(
  'Total Patients'=length(unique(Patient)),
  'Total Samples'=sum(total.samples),
  'Range'=paste(range(end.year-start.year), collapse='-'),
  'Median Years'=paste(median(end.year-start.year),'+/-',round(sd(end.year-start.year),1)),
  'Median Endoscopies'=paste(median(total.endos),'+/-',round(sd(total.endos),1)),
  'Median age'=paste(median(age.diagnosis),'+/-',round(sd(age.diagnosis),1)),
  'gender F:M'=paste(table(gender),collapse=':')
))
, caption='full cohort')
```

```{r echo=F}
cohortSummary = as.data.frame(all.info %>% group_by(Status) %>% summarise(
  'Total Patients'=length(unique(Patient)),
  'Total Samples'=sum(total.samples),
  'Range follow-up (years)'=paste(range(end.year-start.year), collapse='-'),
  'Mean follow-up (years)'=paste(round(mean(end.year-start.year), 1),'+/-',round(sd(end.year-start.year),1)),
  'Range endoscopies'=paste(range(total.endos), collapse='-'),
  'Mean Endoscopies'=paste(round(mean(total.endos), 1),'+/-',round(sd(total.endos),1)),
  'Mean age'=paste(round(mean(age.diagnosis), 1),'+/-',round(sd(age.diagnosis),1)),
  'gender F:M'=paste(table(gender),collapse=':'),
  'Mean Length'=paste(round(mean(C,na.rm=T), 1),'+/-',round(sd(C,na.rm=T),1)),
  'naC'=length(which(is.na(C)))
))

cohortSummary = cbind(cohortSummary,
  rbind(table(subset(all.patient.info, Status == 'NP')$Pathology),
        table(subset(all.patient.info, Status == 'P')$Pathology)))

  rbind(table(subset(patient.info, Status == 'NP')$Pathology),
        table(subset(patient.info, Status == 'P')$Pathology))

  rbind(table(subset(validation.patient.info, Status == 'NP')$Pathology),
        table(subset(validation.patient.info, Status == 'P')$Pathology))

@@

`r pander(t(cohortSummary), caption='Full cohort, split by status')`

@@

```{r echo=F, warning=F, message=F, eval=T, fig.height=12, fig.width=20}

@@

globalCNdf$Status = factor(globalCNdf$Status, levels=sort(unique(globalCNdf$Status), decreasing=T), ordered=T)
globalCN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms'), labeller(Status=n.status, chr=label_value))
cnLegend = get.legend(globalCN.plot)
#globalCN.plot = globalCN.plot + theme(legend.position = 'none')

## Dysplasia only
dys.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology %in% c('LGD','HGD','IMC'))$variable)), ')', sep=''),
              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology %in% c('LGD','HGD','IMC'))$variable)),')', sep=''))

globalnHGDCN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms' & Pathology %in% c('LGD','HGD','IMC')), labeller(Status=dys.status, chr=label_value)) + labs(title='LGD/HGD/IMC Only')

## No HGD
# `%nin%` = Negate(`%in%`)
# 
# nhgd.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology %nin% c('HGD','IMC'))$variable)), ')', sep=''),
#               'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology %nin% c('HGD','IMC'))$variable)),')', sep=''))
# globalnHGDCN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms' & Pathology %nin% c('HGD','IMC')), labeller(Status=nhgd.status, chr=label_value)) + labs(title='Excluding HGD/IMC')

@@

globalCN.plot + theme(legend.position = 'none')

globalnHGDCN.plot + theme(legend.position = 'none')

globalBECN.plot + theme(legend.position = 'none')

# dfpts = subset(globalCNdf, Patient %in% c('57','76') & seg.type != 'arms')
# #dfpts = dfpts %>% group_by(Patient, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
# 
# pt.status <- c('Non-Progressor'=paste('Patient 57 (n=',length(unique(subset(dfpts, Patient == 57)$variable)), ')', sep=''),
#               'Progressor'=paste('Patient 76 (n=',length(unique(subset(dfpts, Patient == 76)$variable)),')', sep=''))
# genomePtCN.plot = cn.mtn.plot(dfpts, labeller(Status = pt.status))
# 
# dfpts = subset(globalCNdf, Patient %in% c('57','76') & Pathology == 'BE' & seg.type != 'arms')
# #dfpts = dfpts %>% group_by(Patient, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
# 
# ptbe.status <- c('Non-Progressor'=paste('Patient 57 (n=',length(unique(subset(dfpts, Patient == 57)$variable)), ')', sep=''),
#               'Progressor'=paste('Patient 76 (n=',length(unique(subset(dfpts, Patient == 76)$variable)),')', sep=''))
# genomePtBECN.plot = cn.mtn.plot(dfpts,  labeller(Status = ptbe.status )) + labs(title='NDBE samples only')
# 
# genomePtCN.plot + theme(legend.position = 'none')
# genomePtBECN.plot + theme(legend.position = 'none')

@@

#a = sample( subset(subset(patient.info, Patient == 57),  Endoscopy.Year == min(Endoscopy.Year) )$Samplename, 1)
#b = sample( subset(subset(patient.info, Patient == 76), Endoscopy.Year == min(Endoscopy.Year) & Pathology %nin% c('HGD','IMC'))$Samplename, 1)

#ind.status <- c('Non-Progressor'='Patient 57 (n=1)','Progressor'='Patient 76 (n=1), excl. HGD/IMC')
#cn.mtn.plot(subset(globalCNdf, variable %in% c(a,b) & seg.type != 'arms'), labeller(Status = ind.status), type='bar') + labs(title="Individual samples")
```

```{r loaddata, message=F, warning=F, echo=F}

file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading", file))
  load(file, verbose=T)
} else {
  stop(paste("No model file", file, "found"))
}

# cx was generated on raw seg values
plot(dysplasia.df[names(labels) ,'cx'], col=labels+1)

# remove old cx...see what new one would look like
dysplasia.df = dysplasia.df[,-grep('cx',colnames(dysplasia.df))]

# general variance per sample after the samples have been mean normalized, ignores segment length
cx = apply(dysplasia.df, 1, function(x) var(x))

x = cbind.data.frame('Status'=labels, 'cx'=cx[names(labels)])
x$Status = factor(x$Status, labels = c('Non-Progressor','Progressor'))
x = merge(x,patient.info[,c('Samplename','Pathology','Patient')],by.x='row.names',by.y='Samplename')

tt = t.test(subset(x, Status == 'Progressor')$cx, subset(x, Status == 'Non-Progressor')$cx)
ggplot(x, aes(Status, cx, fill=Status)) + 
  geom_jitter(width=0.2, color='grey39') + 
  geom_violin(outlier.colour = NA, alpha=0.8) +
  labs(x='', y='Variance', title='Per sample variance', subtitle=paste('p-value=', format.pval(tt$p.value, digits=2))) + theme(legend.position = 'none') +
  plot.theme

rocX = pROC::roc(Status ~ cx, data=x, auc=T, ci=T, of='thresholds')
roc.plot(rocX, title='Per sample variance') + plot.theme

@@

# ch$ymax = ifelse(ch$coef > 0, max(genomePtCN.plot$data$value), min(genomePtCN.plot$data$value))
# chh = chh %>% rowwise() %>% mutate('max'=custom.max(coef, hazards, ymax))
# genomePtCN.plot +  geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F)

# genomePtBECN.plot +  geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F)

@@

```{r insilico_rocs, echo=F, message=F, warning=F, fig.height=8, fig.width=8}

load(file=paste(cache.dir, 'insilico_splits.Rdata', sep='/'), verbose=T)

# m = melt(insilico.rocs, measure.vars = 'auc')
# ggplot(m, aes(variable,value)) + ylim(0,1) +
#   geom_boxplot(fill='mintcream', color='grey55') + geom_jitter(width=0.2, color='darkblue', size=1.5, show.legend = F) + 
#   geom_text(data=subset(m, value %in% range(value)), aes(label=round(value,2)), vjust=0.02, hjust=0.2) +
#   plot.theme + labs(x='',y='AUC',title='AUC', subtitle='20 insilico 20:69 splits')

```{r prediction_cutoff, echo=F, message=F, warning=F, fig.height=7, fig.width=7}
preds = do.call(rbind.data.frame, lapply(pg.samp, function(df) df[c('Status','Prediction')]))
roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds')
roc$model = 'CN Model'

multi.roc.plot(list(roc), colors=brewer.pal(3,'Greys')[3]) + labs(title='Discovery cohort ROC')

insilico.rocs = rbind(insilico.rocs, c(pROC::ci.auc(roc), pROC::coords(roc, 'best')))

insilico.rocs[1:20,'split'] = paste('in-silico', 1:20)
insilico.rocs[21,'split'] = 'discovery'

ggplot(insilico.rocs, aes(split, auc)) + geom_point() + ylim(0,1) +
  geom_errorbar(aes(ymin=auc.ci.min, ymax=auc.ci.max), width=0.2) + 
  coord_flip() + plot.theme + labs(y='AUC', x='', title='AUC per split', subtitle='20 insilico 20:69 splits')

@@

ggplot() + geom_line(data=df, aes(specificity, sensitivity), color='darkblue', size=1.5) + scale_x_reverse() + 
  geom_point(data=thresholds, aes(x=specificity,y=sensitivity,color=label), size=3,shape=8, show.legend = T) +
  geom_label_repel(data=thresholds, aes(x=specificity,y=sensitivity,fill=label,label=
    paste('Threshold ', round(threshold,2), '\n(FPR ', round((1-specificity),2)*100, '%, TPR ', round(sensitivity,2)*100,'%)\nAUC:',round(roc$auc, 2)*100, '%', sep='')),show.legend = F, color='white',nudge_x=0.05) + scale_fill_manual(values=brewer.pal(3,'Dark2')) + scale_color_manual(values=brewer.pal(3,'Dark2')) +
  labs(title='CN Model ROC', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme

@@

@@

# get.auc.at(p53roc)
# get.auc.at(hgdpathroc)
# get.auc.at(pathroc)
# get.auc.at(roc)

@@

rocList = list(roc,pathroc,hgdpathroc,p53roc)

cols = brewer.pal(11, "PRGn")
dodge = position_dodge(width=0.9)
ggplot(melt(subset(m, Measure == 'AUC'), measure.vars='value'), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=Measure)) + 
  geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='black' ) + 
  geom_text(aes(label=paste(round(value,2)*100,'%',sep=''), y=1, vjust=-1 ), position=dodge, color='white')  +
  scale_fill_manual(values=cols, name='') + plot.theme + 
  theme(legend.position = 'none', axis.text.x = element_text(angle=45, hjust=1)) + labs(x='', y='AUC', title='ROC Comparison')

```{r echo=F, fig.height=8, fig.width=6}
roc$model = "Per Sample"

m = rbind.data.frame(
      get.roc.stat(roc),
      get.roc.stat(rocEndoM),get.roc.stat(rocEndoA),
      get.roc.stat(rocPtM), get.roc.stat(rocPtA),
      get.roc.stat(rocFirstEndo), get.roc.stat(rocLastEndo))
m$model = factor(m$model)

cols = brewer.pal(11, 'Paired')
dodge = position_dodge(width=0.9)
mm = melt(subset(m, Measure == 'AUC'), measure.vars='value')
ggplot( arrange(mm, CI.min), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=model)) + 
  geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='black' ) + 
  geom_text(aes(label=paste(round(value,2)*100,'%',sep=''), y=1), position=dodge, color='white')  +
  scale_fill_manual(values=cols, name='') + plot.theme + 
  theme(legend.position = 'none', axis.text.x = element_text(angle=45, hjust=1)) + labs(x='', y='AUC', title='ROC Comparison')

multi.roc.plot(list(roc, rocEndoM, rocEndoA, rocPtM, rocPtA,rocFirstEndo,rocLastEndo), 'ROC Comparisons', palette='Dark2')
```

@@

```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=20}

min.theme = theme(text=element_text(size=10), panel.background=element_blank(), strip.background=element_rect(fill="grey97"), 
                   strip.text = element_text(size=10), axis.text = element_blank(),
                   axis.line=element_line(color='black'), panel.grid.major.y =element_line(color='grey95'), panel.grid.major.x = element_blank(),
                   panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing.x = unit(0, 'lines'), panel.spacing.y = unit(0.1, 'lines')   )

mtn.spatial<-function(df, b, limits, title='', pal=NULL) {
  if (is.null(pal))
    pal = RColorBrewer::brewer.pal(length(limits), 'Set2')

  df = merge(df, b[,c('Samplename','Endoscopy.Year','months.before.final','Block','ogj')], by.x='variable',by.y='Samplename')
  df$ogj = factor(df$ogj, ordered = T)
  df$ogj = factor(df$ogj, levels=rev(levels(df$ogj)), ordered = T)
  #pal = pal[which(limits %in% levels(df$nl))]

  altcolor = c(RColorBrewer::brewer.pal(9, 'Greys')[2],RColorBrewer::brewer.pal(9, 'Greys')[4])
  fills = c()
  for (yr in sort(unique(df$Endoscopy.Year))) {
    ac = (which(sort(unique(df$Endoscopy.Year)) == yr) %% 2) + 1
    fills = c(fills, rep(altcolor[ac], length(which(table(subset(df, Endoscopy.Year == yr)$ogj) > 0))))
  }

  p = ggplot(df, aes(x=chr.length)) + facet_grid(Endoscopy.Year+ogj~chr, scales='free', space='free_x', labeller = labeller(.multi_line=F)) +
      geom_rect(aes(xmin=start, xmax=end, ymin=0, ymax=mean.value, fill=ogj), show.legend=T) + 
      scale_fill_manual(values=pal, limits=limits,  name='OGJ Dist.') +
      labs(title=title, x='') + min.theme

  legend = get.legend(p)
  p = p + theme(legend.position = 'none')

  g = ggplot_gtable(ggplot_build(p))
  strips = grep('strip-r', g$layout$name)

  for (n in 1:length(strips)) {
    i = strips[n]
    j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
    g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[n]
  }

  grid.arrange(g)
  return(list('grob'=g, 'legend'=legend))
  #grid::grid.draw(g)  
  #return(g)
}

back$ogj = back$nl
back$ogj[which(back$nl > 0 & back$nl < 3)] = 1 
back$ogj[which(back$nl > 3 & back$nl < 7)] = 2 
back$ogj[which(back$nl >= 7)] = 3

ogjLimits = levels(factor(back$ogj, ordered = T))

p1 = mtn.spatial(subset(globalCNdf, Patient == 48 & seg.type != 'arms'), back, ogjLimits, 'Patient 48 (P)')
p2 = mtn.spatial(subset(globalCNdf, Patient == 20 & seg.type != 'arms'), back, ogjLimits,'Patient 20 (P)' )
p3 = mtn.spatial(subset(globalCNdf, Patient == 60 & seg.type != 'arms'), back, ogjLimits,'Patient 60 (P)' )

p4 = mtn.spatial(subset(globalCNdf, Patient == 3 & seg.type != 'arms'), back,ogjLimits,'Patient 3 (NP)' )
p5 = mtn.spatial(subset(globalCNdf, Patient == 15 & seg.type != 'arms'), back,ogjLimits,'Patient 15 (NP)' )
p6 = mtn.spatial(subset(globalCNdf, Patient == 51 & seg.type != 'arms'), back,ogjLimits,'Patient 51 (NP)' )

back$ogj = NULL

```{r, echo=F, fig.width=2, fig.height=4}

grid.arrange(p1$legend)

@@

```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=12}
riskPath = rbind( cbind.data.frame( table(subset(back, Status == 'P', select=c('Pathology','Risk'))), 'Status'='Progressor' ), 
                  cbind.data.frame( table(subset(back, Status == 'NP', select=c('Pathology','Risk'))), 'Status'='Non-Progressor' ))

@@

riskPath = merge(riskPath, pathTotal)
riskPath$ratio = with(riskPath, round(Freq/nPath, 3))
riskPath$Risk = factor(riskPath$Risk, levels=c('Low','Moderate','High'), ordered = T)

p53Path = cbind.data.frame(table(subset(back, Status == 'P', select=c('Pathology','p53.Status'))), 'Status'='p53 IHC')
p53Path = merge( subset(back, Status == 'P' & !is.na(p53.Status)) %>% group_by(Pathology) %>% summarise( 'nPath'=length(Pathology)), p53Path)
#p53Path$Risk = with(p53Path, ifelse(p53.Status == 1, 'High','Low'))
p53Path$ratio = with(p53Path, round(Freq/nPath, 2))

@@

ggplot(riskPath, aes(Pathology, ratio*100)) + geom_bar(aes(group=Risk, fill=Risk),stat='identity', position=position_stack()) + 
  scale_fill_manual(values=riskCols) + scale_color_manual(values=c('white','black','white')) +
  geom_text(data=pathTotal, aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101), position=position_stack(), size=4) +
  geom_text(data=subset(riskPath, ratio>0), aes(group=Risk,label=paste(ratio*100,'%',sep=''), color=Risk), position=position_stack(vjust = 0.5), size=5, show.legend = F) +
  facet_grid(~Status) + plot.theme + labs(title='Samples predicted by pathology', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')

@@

```{r p53IHC, echo=F, warning=F, message=F, fig.height=8, fig.width=8}

#p53Path$Risk[p53Path$Risk == 'High'] = 'Aberrant'
#p53Path$Risk[p53Path$Risk == 'Low'] = 'Normal'

#p53Path$p53.Status = factor(p53Path$p53.Status, levels=c('Aberrant', 'Normal'), ordered = F)

p53Path$p53.Status = factor(p53Path$p53.Status, ordered = T)
p53Path = arrange(p53Path, Pathology, p53.Status)
col = brewer.pal(11, "PRGn")[c(9,3)]
ggplot(p53Path, aes(Pathology, ratio*100, group=Pathology, fill=p53.Status)) + geom_bar(stat='identity', position = position_stack()) +
  scale_fill_manual(values=col, name='p53 IHC', labels=c('Normal','Aberrant')) + 
  geom_text(aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101)) +
  geom_text(aes(label=paste(ratio*100,'%',sep='')), position=position_stack(vjust = 0.5), size=5) +
  plot.theme + labs(title='Samples predicted by aberrant p53 IHC', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')

@@

```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=5 }
back$nl = factor(back$nl, ordered=T)
back$Risk = factor(back$Risk, levels=c('Low','Moderate','High'), ordered = T)
back$Patient = factor(as.numeric(back$Patient), ordered = T)

p = pred.tiles(subset(back, Patient %in% c(6)), ncol=1, cols=myPal, probs=T) + labs(title='')
tilesLegend = get.legend(p)
p + theme(legend.position = 'none')

pred.tiles(subset(back, Patient %in% c(6)), ncol=1, cols=myPal, probs=T, OR=F) + labs(title='') + theme(legend.position = 'none')
pred.tiles(subset(back, Patient %in% c(48)), ncol=1, cols=myPal, probs=T) + labs(title='') + theme(legend.position = 'none')
pred.tiles(subset(back, Patient %in% c(33)), ncol=1, cols=myPal, probs=T) + labs(title='')+ theme(legend.position = 'none')
pred.tiles(subset(back, Patient %in% c(89)), ncol=1, cols=myPal, probs=T) + labs(title='')+ theme(legend.position = 'none')
```

```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=15 }
grid.arrange(
  pred.tiles(subset(back, Patient %in% c(13,20,34,56,60,48) & Status == 'P'), ncol=1, cols=myPal, probs=T) + labs(title='Progressors') + theme(legend.position = 'none'),
  pred.tiles(subset(back, Patient %in% c(15,16,21,3,45,19) & Status == 'NP'), ncol=1, cols=myPal, probs=T) + labs(title='Non-Progressors') + theme(legend.position = 'none'),
ncol=2)

@@

@@

roc$model = 'Discovery'
multi.roc.plot(list(roc, vroc), title='ROC', palette='Dark2') 
```
